---
title: 'Cyclistic Bike Share: Analytics'
author: "David Jackson"
date: "8/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,error = FALSE)
library(tidyverse)
library(janitor)
library(lubridate)
rm(list=ls())
```
```{r}
##
## Import previous 12 months data
##
df1 <- read.csv("./Data/t1.csv")
df2 <- read.csv("./Data/t2.csv")
df3 <- read.csv("./Data/t3.csv")
df4 <- read.csv("./Data/t4.csv")
df5 <- read.csv("./Data/t5.csv")
df6 <- read.csv("./Data/t6.csv")
df7 <- read.csv("./Data/t7.csv")
df8 <- read.csv("./Data/t8.csv")
df9 <- read.csv("./Data/t9.csv")
df10 <- read.csv("./Data/t10.csv")
df11 <- read.csv("./Data/t11.csv")
df12 <- read.csv("./Data/t12.csv")
```
```{r}
##
## Combine 12 data.frames into One (1) data.frame
##
bike_rides <- rbind(df1,df2,df3,df4,df5,df6,df7,df8,df9,df10,df11,df12)
bike_rides <- janitor ::remove_empty(bike_rides,which = c("cols"))
bike_rides <- janitor::remove_empty(bike_rides,which = c("rows"))

```
```{r}
##
## Convert Data/Time stamp to Date/Time ...
##
bike_rides$Ymd <- as.Date(bike_rides$started_at)
bike_rides$started_at <- lubridate::ymd_hms(bike_rides$started_at)
bike_rides$ended_at <- lubridate::ymd_hms(bike_rides$ended_at)

bike_rides$start_hour <- lubridate::hour(bike_rides$started_at)
bike_rides$end_hour <- lubridate::hour(bike_rides$ended_at)
```

```{r}

bike_rides$Hours <- difftime(bike_rides$ended_at,bike_rides$started_at,units = c("hours"))

bike_rides$Minutes <- difftime(bike_rides$ended_at,bike_rides$started_at,units = c("mins"))


df <- bike_rides %>% filter(Hours >0) %>% drop_na() %>%
  select(-ride_id,-end_station_name,-end_station_id,
        -end_station_name )
```

## Data Wrangling Change Log

August 6, 2021:

* Total Rows: 3,489,749
* I raised the following three concerns (about the data) with Lily Moreno(The director of marketing): 
(1) 10,552 rows  with negative trip durations,
(2) 122,175 rows with missing Starting Station names (and ID),(3) As many as 1,291 stations  with two different start_staition_id.

* I was advised by Lily Moreno that we could ignore rows with trip duraton <=0.
We could also ignore rows with missing start_station_name, and also ignore stations_id. 
* And should focus on doing using the start_station_name to prefrom aggregate functions on date, start_station_name, member_casual and rideable_type.
* Cleaning reducted rows to 3,343,689

